---
title: "US Data Overview"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include = FALSE}
library(tidyverse)
library(httr)
library(plotly)
library(kableExtra)


knitr::opts_chunk$set(
 message = FALSE, warning = FALSE
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Raw Data:

```{r}
load_df =
GET("https://chronicdata.cdc.gov/resource/fqb7-mgjf.csv",
    query = list("$limit" = 50000)) %>% 
  content("parsed") %>% 
  janitor::clean_names()


tidy_df = 
  load_df %>% 
  filter(is.na(data_value_footnote)) %>% 
  dplyr::select(-class, -topic, -question, -data_value_unit, -data_value_type, -data_source, -data_value_footnote_symbol, -data_value_footnote, -question_id, -class_id, -topic_id, -states, -counties)

head(tidy_df) 

```

## Important Variables

* `year`: 2011 - 2019
* `locationabbr`: state abbreviation, 56 states
* `locationdesc`: state name, 56 states
* `response`:Normal Weight (BMI 18.5-24.9); Obese (BMI 30.0 - 99.8); Overweight (BMI 25.0-29.9); Underweight (BMI 12.0-18.4); Underweight (bmi 12.0-18.4);Normal Weight (bmi 18.5-24.9); Obese (bmi 30.0 - 99.8); Overweight (bmi 25.0-29.9)
* `break_out_category`: 
  * Race/Ethnicity; Age Group; Household Income; Education Attained; Gender; Overall  
* `break_out`: 
  * Male; College graduate; White, non-Hispanic; `$35,000-$49,999`; `$25,000-$34,999`; `$15,000-$24,999`; 35-44; 25-34; 18-24; 55-64; (Other) ; Overall; Some post-H.S.
* `sample_size`
* `break_out_category_id`: gender(CAT2), age(CAT3), race/ethnicity(CAT4), education(CAT5)


## Different Education Attained
```{r}
education_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT5") 
head(education_df) 
```

## Different Race/Ethnicity
```{r}
race_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT4") 
head(race_df) 
```

## Different Age
```{r}
age_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT3") 
head(age_df) 
```

## Different Gender
```{r}
gender_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT2") 
head(gender_df) 
```


## Time Trends

```{r tidy}
# total population in each state in each year
df.tot = load_df %>%
  group_by(year,locationdesc) %>%
  summarise(state_sum = sum(sample_size))

# total 'obese' and 'overweight' amount in each state in each year
df.overweight = 
  load_df %>%
  filter(response %in% c("Obese (BMI 30.0 - 99.8)","Overweight (BMI 25.0-29.9)",
                         "Obese (bmi 30.0 - 99.8)","Overweight (bmi 25.0-29.9)")) %>%
  group_by(year,locationdesc) %>%
  summarise(state_overweight_sum = sum(sample_size))


df.rate = left_join(df.tot, df.overweight,
                   by = c("year" , "locationdesc")) %>%
  mutate(over_rate = state_overweight_sum/state_sum)

```

```{r  out.width = "120%"}
# Use library(plotly) to visualizing the US obesity and overweight rate geographically (on a map), including the average amount of obesity and overweight from 2011 to 2019.
df.rate1 =  df.rate %>%
  group_by(locationdesc)  %>%
  mutate(mean_amount = mean(state_overweight_sum))

df.rate1 = df.rate1[-which(df.rate1$locationdesc %in% c("All States and DC (median) **","All States, DC and Territories (median) **","Virgin Islands") ),]
df.rate1$region = tolower(df.rate1$locationdesc)

states = map_data("state")
# merge the datasets
states = merge(states, df.rate1, by = "region", all.x = T)

##### US states obesity time trend: scatter plot with line
ggplot(data = df.rate[-which(df.rate$locationdesc %in% c("All States and DC (median) **","All States, DC and Territories (median) **","Virgin Islands") ),], 
       aes(year, y = over_rate)) +
  geom_line(aes(color = locationdesc)) +
  geom_point(aes(color = locationdesc),size = 1.5) +
  facet_wrap(~ locationdesc) +
  theme(legend.position = "None",
        #text = element_text(family='Kai'),
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_continuous(breaks = seq(2011,2019,1)) +
  xlab("Year") + ylab("Obesity and overweight rate") +
  ggtitle("Obesity and overweight rate trend for all the states")
```

It is clear that **all states** have as **increasing** trend of obese and overweight amount from 2011 to 2019.


## The state with the highest average obesity and overweight throughout 9 years 
```{r}
df.highest = df.rate %>% 
  group_by(locationdesc) %>%
  summarise(mean_overweight = mean(state_overweight_sum)) %>%
  filter(mean_overweight == max(mean_overweight)) 


kable(df.highest, row.names = T, escape = F) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

We find the state `Nebraska` has the highest average `obese` and `overweight` amount per year among all the states.





