---
title: "Regression Analysis"
output:   
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    
---

```{r,include=FALSE,message=FALSE,echo=FALSE}
library(tidyverse)
library(ggplot2)
library(broom)
library(httr)
library(modelr)
library(readr)
library(mgcv)
set.seed(1)

knitr::opts_chunk$set(
echo = FALSE,
 fig.width = 7, 
 fig.height = 5,
 fig.asp = 0.6,
 out.width = "90%")
theme_set(theme_bw() + 
          theme(legend.position = "bottom",
                legend.title = element_blank(),
                plot.title = element_text(hjust = 0.5, size = 15),
                plot.subtitle = element_text(hjust = 0.5, size = 12)))

ob_df =
  read_csv("./data/ObesityDataSet_raw_and_data_sinthetic.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    n_obeyesdad = case_when(
      n_obeyesdad == "Normal_Weight" ~ 21.7,
      n_obeyesdad == "Overweight_Level_I" ~ 25,
      n_obeyesdad == "Overweight_Level_II" ~ 30,
      n_obeyesdad == "Obesity_Type_I" ~ 32.5,
      n_obeyesdad == "Insufficient_Weight" ~ 18.5,
      n_obeyesdad == "Obesity_Type_II" ~ 37.5,
      n_obeyesdad == "Obesity_Type_III" ~ 40,
    )
  ) %>% 
  rename(bmi = n_obeyesdad,
         fml_his = family_history_with_overweight) %>% 
  dplyr::select(bmi, gender, age, fml_his, favc, calc, faf) %>% 
  drop_na() %>% 
  mutate(gender = factor(gender),
         fml_his = factor(fml_his),
         favc = factor(favc),
         calc = factor(calc))
```


## Data

The data about the estimation of **_obesity levels_** in people from the countries of Mexico, Peru and Colombia, with ages between 14 and 61 and **_diverse eating habits and physical conditions_**.

This data was collected using a web platform with a survey where anonymous users answered each question, then the information was processed obtaining _17 attributes and 2111 records_. 

<br/>

**Goals**

There are many factors associated with overweight or obesity conditions, including **_congenital factors_** gender, age, and family history with overweight family members. Also, there are **_acquired ones_** also play important roles in weight status, including high caloric food consumption, alcohol consumption and little physical activities. 

To better understand **_overweight status_**, we conduct 3 models, including 6 predictors and their interaction terms, to explore congenital and acquired factors on weight conditions.

<br/>

**Adjustments**

To conduct linear regression, the outcome should be numeric variables, so we convert variable recording obesity condition, such as `Normal_weight` and `Obesity_Type_I` to BMI `21.7` and `32.5` according to [BMI and weights relationships of CDC guildlines](https://www.cdc.gov/obesity/adult/defining.html). Since each weight condition has a BMI range like "25-30", we assume that people with a weight status has an average BMI, that is 27.5. 

<br/>

**Variables**

Head of the dataset: 

`r head(ob_df) %>% knitr::kable()`

- `favc`: Frequent consumption of high caloric food
- `calc`: Consumption of alcohol
- `faf`: Physical activity frequency
- `fml_his`: Family history with overweigh

<br/>

## BMI Distribution

```{r, warning=FALSE, message=FALSE, collapse=FALSE}
ggplot(aes(x=bmi),data = ob_df)+
  geom_density(color="darkblue", fill="lightblue") +
  geom_vline(data=ob_df, aes(xintercept=mean(bmi)),
             linetype="dashed") +
  labs(   # add lables
    title = "BMI distribution",
    x = "BMI",
    y = "Frequency",
    caption = "Dataset of 2111 records, with continuity correction"
  ) 
```

Since we made a **_continuity correction_** to the BMI, there will be gaps between values, like the gap between BMI value 25 and 30. However, we can observe the BMI distribution is normal to some extent.

<br/>

## Three Linear Models

```{r, warning=FALSE, message=FALSE, collapse=FALSE}
reg_bmi1 = lm(bmi ~ gender + age + fml_his + favc + calc + faf, data = ob_df)
reg_bmi2 = lm(bmi ~ gender + age + fml_his + favc * calc * faf, data = ob_df)
reg_bmi3 = lm(bmi ~ gender + age + fml_his, data = ob_df)
```

$Model~1:bmi_{i}=\beta_{0}+\beta_{gender}+\beta_{age}+\beta_{fml-his}+\beta_{favc}+\beta_{calc}+\beta_{faf}+\varepsilon_{i}$

$Model~2:bmi_{i}=\beta_{0}+\beta_{gender}+\beta_{age}+\beta_{fml-his}+\beta_{favc}+\beta_{calc}+\beta_{faf}+\beta_{favc*calc}+\beta_{favc*faf}+\beta_{faf*calc}+\beta_{favc*calc*faf}+\varepsilon_{i}$

$Model~3:bmi_{i}=\beta_{0}+\beta_{gender}+\beta_{age}+\beta_{fml-his}+\varepsilon_{i}$

We have 3 models. <br/>
First is with all factors, 3 congenital and 3 acquired influncing factors. <br>
Second model considers interation between caloric food consumption, alcohol consumption and physical activities' frequencies. <br>
The last model focus only on congenital factors, which is nested in the first model.

<br/>

### First Model with 6 Predictors

```{r, fig.width=11, fig.height=8, warning=FALSE, message=FALSE, collapse=FALSE}
par(mfrow=c(2,2))
plot(reg_bmi1)
```

Based on the first model, we have 4 plots, and give brief analysis on them.

* Residual vs Fitted: residual values bounce around 0, form a horizontal line over there, which shows that the error variance of the data is fairly equal.

* Quantile-Quantile Plot: we can observe a straight line in the first half of the plot, indicating the residuals are normal. There present outliers in the latter half though.

* Scale-Location: There is a horizontal line, showing that the variance of distribution is equal.

* Residual vs Leverage: Though not all outliers are influential, there are some outliers beneath the Cook's distance.

* Plot Residual vs Fitted and Scale-Location may seems discrete and weired since we used continuity correction, but it dosen't influence much.

<br/>

### Second Model with Interactions
```{r, warning=FALSE, message=FALSE, collapse=FALSE}
reg_bmi2 %>% 
  broom::tidy() %>% 
  drop_na() %>% 
  knitr::kable()
```
 
From the table showing coefficients of the second model, we observe there are **_several factors are influencial along with some interactions_**, including `gender`, `age`, `fml_his`, `favc`, `favc*calc`.

All congenital factors matters, while only high caloric food consumption and its interation with alcohol consumption lead to overweight conditions.
 
<br/>
 
## Comparing three models

<br/>

### Cross Validation for comparisons

**Adjustment**

The cross validation requires all predictors to be numeric variables, so we transform variables `gender`, `fml_his`, `favc`, `calc` to numeric ones. Since we need to know `rmse` (root mean square error) distributions of models, and basic trends of BMI influenced by variables, the addjustment is acceptable.

```{r, warning=FALSE, message=FALSE, collapse=FALSE}
ob_dbl =
  ob_df %>% 
  mutate(gender = recode(gender, "Female" = 0, "Male" = 1),
         fml_his = recode(fml_his, "yes" = 1, "no" = 0),
         favc = recode(favc, "yes" = 1, "no" = 0),
         calc = recode(calc, "no" = 0, "Sometimes" = 1, "Frequently" = 2, "Always" = 3))

reg_dbl1 = lm(bmi ~ gender + age + fml_his + favc + calc + faf, data = ob_df)
reg_dbl2 = lm(bmi ~ gender + age + fml_his + favc * calc * faf, data = ob_df)
reg_dbl3 = lm(bmi ~ gender + age + fml_his, data = ob_df)
```

```{r, warning=FALSE, message=FALSE, collapse=FALSE}
cv_df =
  crossv_mc(ob_dbl, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(
    reg_dbl1  = map(.x = train, ~lm(bmi ~ gender + age + fml_his + favc + calc + faf, data = .x)),
    reg_dbl2  = map(.x = train, ~lm(bmi ~ gender + age + fml_his + favc * calc * faf, data = .x)),
    reg_dbl3  = map(.x = train, ~lm(bmi ~ gender + age + fml_his, data = .x))) %>% 
  mutate(
    rmse_model1 = map2_dbl(.x = reg_dbl1, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model2 = map2_dbl(.x = reg_dbl2, .y = test, ~rmse(model = .x, data = .y)),
    rmse_model3 = map2_dbl(.x = reg_dbl3, .y = test, ~rmse(model = .x, data = .y)))
    
```

```{r, warning=FALSE, message=FALSE, collapse=FALSE}
cv_df %>% 
  dplyr::select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
```

In the plot showing RMSE, we see the model 1 and model 2 are with the smallest RMSE, showing that those two models fit well. For the third model only focus on congenital factors, it confirms the inference in the coefficients table that **_acquired factors also matter_**.

<br/>

### ANOVA test for comparisons

```{r, warning=FALSE, message=FALSE, collapse=FALSE}
anova(reg_bmi3, reg_bmi1)
```

Finally, we use the **ANOVA** test to compare significance of nested model 3 in model 1. Since the P value of the test is _2.2e-16_, way smaller than 0.05 significance. So we may have the conclusion that model 1, the large model containing acquired influential factors, fits best.

<br/>

## Further Exploration

```{r, warning=FALSE, message=FALSE, collapse=FALSE}
ob_df %>% 
  lme4::lmer(bmi ~ gender + fml_his + (1 + fml_his | favc), data = .) %>% 
  broom.mixed::tidy() %>% 
  knitr::kable()
```

Also, we can use a **_mixed model_** for further exploration, since there are many factor levels and models fitting with main effects or interactions for each is not appropriate. Though out of the scope of our course, we may use the mixed model to attain random intercepts and slopes of each factors.

<br/>

<br/>
