---
title: "Visualizations"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(httr)
library(ggplot2)
library(broom)
library(shiny)


knitr::opts_chunk$set(
 echo = FALSE,
 fig.width = 7, 
 fig.height = 5,
 fig.asp = 0.6,
 out.width = "60%")
theme_set(theme_bw() + 
          theme(legend.position = "bottom",
                legend.title = element_blank(),
                plot.title = element_text(hjust = 0.5, size = 15),
                plot.subtitle = element_text(hjust = 0.5, size = 12)))

load_df =
GET("https://chronicdata.cdc.gov/resource/fqb7-mgjf.csv",
    query = list("$limit" = 50000)) %>% 
  content("parsed") %>% 
  janitor::clean_names()


ne_df = 
  load_df %>% 
  filter(is.na(data_value_footnote)) %>% 
  select(-class, -topic, -question, -data_value_unit, -data_value_type, -data_source, -data_value_footnote_symbol, -data_value_footnote, -question_id, -class_id, -topic_id, -states, -counties) 
```

```{r, echo=FALSE}
shinyApp(
options(shiny.sanitize.errors = TRUE),
ui = shinyUI(fluidPage(titlePanel(""),
                                    # sidebar style
                                         sidebarLayout(sidebarPanel(
                                           #selectInput function
                                           selectInput("location","Please Select locationaddr:",unique(ne_df$locationdesc))),
                                           mainPanel(
                                             plotlyOutput("plot1"),
                                             br(),
                                             br(),
                                             plotlyOutput("plot2"),
                                             br(),
                                             br(),
                                             plotlyOutput("plot3"),
                                             br(),
                                             br(),
                                             plotlyOutput("plot4"),
                                             br(),
                                             br(),
                                             plotlyOutput("plot5")
                                           )))),
server = function(input, output, session) {
  
  #choose 
  Pass <- reactive({
    ne_df[ne_df$locationdesc==input$location,]
  })
  output$plot1 <- renderPlotly({
    time_df = 
  Pass() %>% 
  filter(response_id %in% c("RESP039", "RESP040") ) %>% 
  select(year, response_id, sample_size) %>% 
  group_by(year, response_id) %>% 
  mutate(count_obese = sum(sample_size)) %>% 
  select(-sample_size) %>% 
  distinct()

total_df = 
  Pass() %>%
  select(year, response_id, sample_size) %>% 
  group_by(year) %>% 
  mutate(count_total = sum(sample_size)) %>% 
  select(-sample_size) %>% 
  filter(response_id %in% c("RESP039", "RESP040") ) %>% 
  distinct() 

rate_time = 
  left_join(time_df, total_df,by = c("year" , "response_id")) %>% 
  mutate(
    prevalence = count_obese / count_total,
    rate_lable = case_when(
      response_id == "RESP039" ~ "Obese (BMI 30.0 - 99.8)",
      response_id == "RESP040" ~ "Overweight (BMI 25.0-29.9)"
    )
    ) %>%
  mutate(prevalence = round(prevalence, digits = 4)) 

rate_time <- rate_time %>%
  group_by(rate_lable) %>% 
  mutate(text_label = str_c("Year:", year, "\nPrevalence:", prevalence)) 

# plot 1: obese and overweight by states: 
  plot_ly( rate_time,
    x = ~year, 
    y = ~prevalence, 
    #type = 'scatter', 
    #mode = "lines+markers", 
    textposition = "bottom right",  
    color = ~rate_lable, 
    colors = c('#FFD700', '#FF8C00'),
    text = ~text_label, 
    line = list(simplyfy = F),
    alpha = 0.8
    ) %>% 
  layout(
    legend = list(x = 0.5, y = 0.15),
    title = "Obesity rate trend over time",
         xaxis = list(title = "Year",
                      zeroline = TRUE),
         yaxis = list(title = "Prevalence",
                      zeroline = TRUE),
     updatemenus = list(
      list(
        x = 0.7,
        y = 0.4,
        buttons = list(

          list(method = "restyle",
               args = list("type", "scatter"),
               label = "Line"),

          list(method = "restyle",
               args = list("type", "bar"),
               label = "bar")))
    )
    ) 

  })
  
 #Difference Among Race Groups:
  
  output$plot2 <-renderPlotly({
    ne_race_df = 
  Pass() %>% 
  filter(break_out_category_id == "CAT4") %>% 
  select(year, locationabbr, locationdesc, response, break_out, sample_size, response_id) %>% 
  group_by(year, break_out) %>% 
  mutate(count_race = sum(sample_size)) %>% 
  filter(response_id %in% c("RESP039") ) %>% 
  mutate(race_pre = round(sample_size / count_race, digit = 4))


fig_race_2 = 
  ne_race_df %>% 
  filter(break_out != "Other, non-Hispanic") %>% 
  mutate(text_label = str_c("Race/Ethnicity:", break_out, "\nPrevalence:", race_pre)) 


  plot_ly(fig_race_2, x = ~year , y = ~race_pre,  type = "bar", color = ~break_out, text = ~text_label, alpha = 0.7) %>% 
  layout(
         
         title = "Difference Among Race Groups",  
         legend = list(font = list(size = 8),  x = 0.5, y = 0.05),
         xaxis = list(title = "Year",
                      zeroline = TRUE,
                            rangeselector = list(
        buttons = list(
          list(
            count = 3,
            label = "3 year",
            step = "year",
            stepmode = "backward"),
          list(
            count = 6,
            label = "6 year",
            step = "year",
            stepmode = "backward"),
          list(
            count = 1,
            label = "YTD",
            step = "year",
            stepmode = "todate"),
          list(step = "all"))),

      rangeslider = list(type = "date")
                      ),
                      
         
         yaxis = list(title = "Prevalence",
                      zeroline = TRUE)) 



  })
  
  
# Difference Between Genders:
  
 output$plot3 <- renderPlotly({
   ne_gender_df = 
  Pass() %>% 
  filter(break_out_category_id == "CAT2") %>% 
  select(year, locationabbr, locationdesc, response, break_out, sample_size, response_id) %>% 
  group_by(year, break_out) %>% 
  mutate(count_gender = sum(sample_size)) %>% 
  filter(response_id %in% c("RESP039") ) %>% 
  mutate(gender_pre = round(sample_size / count_gender, digit = 4))

ay = list(
  tickfont = list(color = "#767B84"),
  overlaying = "y",
  side = "right",
  title = "Prevalence",
  automargin = TRUE
)
ne_gender_df <- ne_gender_df %>%
  dplyr::group_by(break_out) %>% 
  mutate(
    text_label_1 = str_c("gender:", break_out, "\nCount:", sample_size),
    text_label_2 = str_c("gender:", break_out, "\nPrevalence:", gender_pre)     
    ) 
  plot_ly( data=ne_gender_df)%>%add_bars(x = ~year, 
    y = ~sample_size,  
    type = "bar", 
    color = ~break_out, 
    colors = c("#F1948A", "#85C1E9"),
    text = ~text_label_1, 
    alpha = 0.7) %>% 
    add_lines(x = ~year, y = ~gender_pre, color = ~break_out, text = ~text_label_2, yaxis = "y2") %>% 
  layout(
         title = "Difference Between Genders",
         xaxis = list(title = "year",
                      zeroline = TRUE),
         yaxis = list(title = "Count",
                      zeroline = TRUE),
         yaxis2 = ay,
         legend = list(  
                     x = 0.7,
                     y = 0.1)
         )

 })
 
 # Education Attained group:
 output$plot4 <- renderPlotly({
  ne_edu_df = 
  Pass() %>% 
  filter(break_out_category_id == "CAT5") %>% 
  select(year, locationabbr, locationdesc, response, break_out, sample_size, response_id) %>% 
  group_by(year, break_out) %>% 
  mutate(count_edu = sum(sample_size)) %>% 
  filter(response_id %in% c("RESP039") ) %>% 
  mutate(edu_pre = round(sample_size / count_edu, digit = 4))

edu_rate = 
  ne_edu_df %>% 
  select(break_out, count_edu, year) %>% 
  distinct() %>% 
  group_by(break_out) %>% 
  mutate(sum_edu = round(sum(count_edu/9), digits = 0))

fig_edu_1 = 
  edu_rate %>% 
  plot_ly(labels = ~break_out, values = ~sum_edu, 
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#483C32'),
        hoverinfo = 'text',
        domain = list(row = 0, column = 1),
        text = ~paste('Education Attained:', break_out, ' \nNumber', count_edu),
        marker = list(colors = c('#cd7eaf', '#f3cec9','#6f4d96' ,'#e7a4b6' ),
                      line = list(color = '#FFFFFF', width = 1)),
        showlegend = F
        ) %>% add_pie(hole = 0.5)

fig_edu_2 = 
  ne_edu_df %>% 
  mutate(text_label = str_c("Education Attained:", break_out, "\nPrevalence:", edu_pre)) %>% 
  plot_ly( 
    x = ~year , 
    y = ~edu_pre, 
     domain = list(row = 0, column = 0),
    type = "bar", 
    color = ~break_out, 
    colors = c('#6f4d96', '#e7a4b6', '#f3cec9','#cd7eaf' ),
    text = ~text_label,
    alpha = 0.7) %>% 
  layout(
         title = "Difference between Education Attained group",  
         legend = list( showlegend = F),
         xaxis = list(title = "Education Attained",
                      zeroline = TRUE),
         yaxis = list(title = "Prevalence",
                      zeroline = TRUE)) 

subplot(fig_edu_2 , fig_edu_1) %>%  layout(legend = list(x = 0.3, y = 0.1, font = list(size = 8)), grid = list(rows = 1, columns = 2))
    })
 
 
# "Obesity rate trend over time Income: 
 output$plot5 <- renderPlotly({
   ne_income_df = 
  Pass() %>% 
  filter(break_out_category_id == "CAT6") %>% 
  select(year, locationabbr, locationdesc, response, break_out, sample_size, response_id) %>% 
  group_by(year, break_out) %>% 
  mutate(count_income = sum(sample_size)) %>% 
  filter(response_id %in% c("RESP039", "RESP040")) %>% 
  mutate(
    income_pre = round(sample_size / count_income, digit = 4),
    rate_lable = case_when(
      response_id == "RESP039" ~ "Obese (BMI 30.0 - 99.8)",
      response_id == "RESP040" ~ "Overweight (BMI 25.0-29.9)")
    
    )
  


   ggplot(data=ne_income_df,aes(x = year, y = income_pre, color = rate_lable)) +
   geom_line(size = 1) +
   scale_color_manual(values=c("#20B2AA", "#228B22")) +
   theme_bw() +
   theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45,  hjust = 1)) +
   facet_wrap(~break_out) +
   theme(legend.position = "bottom") +
    labs(title = "Obesity rate trend over time",
       subtitle = "2011 - 2019.",
       x = "Year", 
       y = "Prevalence") +
  scale_x_continuous(breaks = seq(2011,2019,3)) 
   
 })
},
options=list(
    width="120%", height=2400)
)





