---
title: "Data tidy"
output: html_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(httr)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Add data and tidy it

```{r}
load_df =
GET("https://chronicdata.cdc.gov/resource/fqb7-mgjf.csv",
    query = list("$limit" = 50000)) %>% 
  content("parsed") %>% 
  janitor::clean_names()


tidy_df = 
  load_df %>% 
  filter(is.na(data_value_footnote)) %>% 
  select(-class, -topic, -question, -data_value_unit, -data_value_type, -data_source, -data_value_footnote_symbol, -data_value_footnote, -question_id, -class_id, -topic_id, -states, -counties)
```
### Variable : notation
year: 2011 - 2019
locationabbr: state abbreviation, 56 states
locationdesc : state name, 56 states
response :Normal Weight (BMI 18.5-24.9); Obese (BMI 30.0 - 99.8)
          Overweight (BMI 25.0-29.9); Underweight (BMI 12.0-18.4)
          Underweight (bmi 12.0-18.4);Normal Weight (bmi 18.5-24.9)
          Obese (bmi 30.0 - 99.8); Overweight (bmi 25.0-29.9)
break_out_category: Race/Ethnicity; Age Group; Household Income
                    Education Attained; Gender; Overall  
break_out:Male; College graduate;White, non-Hispanic;
          `$35,000-$49,999`; `$25,000-$34,999`; `$15,000-$24,999`;
          35-44; 25-34; 18-24; 55-64; 
          (Other) ; Overall; Some post-H.S.
sample_size
data_value: crude prevalence, unit %, such as 14.7 means 14.7%
confidence_limit_low
confidence_limit_high
display_order: I don't know
geo_location
break_out_category_id: 6 types
breakout_id: such as SEX1, SEX2
location_id: WTF!!! out-of-order!!!! hope it useless
response_id: 4 types


## Data of different household income
```{r}
income_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT6") %>%
  select(year, locationabbr, locationdesc, response, break_out, sample_size, data_value, confidence_limit_low, confidence_limit_high, response_id) %>% 
  group_by(break_out, response) %>%
  mutate(
    sample_sum  = sum(sample_size),
    average_value = weighted.mean(data_value, sample_size)
  ) 
```

## Take NY 2019 for example
```{r}
ny_2019 = 
  income_df %>% 
  filter( year == 2019, locationabbr == "NY") 

ny_2019 %>% 
   mutate(text_label = str_c("Income: ", break_out, "\nBMI: ", response, "\nCrude prevalence: ", average_value)) %>% 
   plot_ly(
      type = "scatter3d", 
      x =  ~break_out, 
      y =  ~response, 
      z = ~average_value,
      text = ~text_label,
      color = ~average_value,
      mode = "markers" 
      )
```
This plot is so disgusting!!!!

## Data of different Education Attained
```{r}
education_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT5") 
```

## Data of different Race/Ethnicity
```{r}
race_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT4") 
```

## Data of different Age
```{r}
age_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT3") 
```

## Data of different Gender
```{r}
gender_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT2") 
```

## Data of nonuse
```{r}
nonuse_df = 
  tidy_df %>% 
  filter(break_out_category_id == "CAT1") 
```
This break_out might be unanswered. Since the dim is 1978 x 16 instead of 38172 x 16