---
title: "Nebraska"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    
---

```{r, echo = FALSE, warning= FALSE, message=FALSE}
library(tidyverse)
library(plotly)
library(httr)
library(ggplot2)
library(broom)



knitr::opts_chunk$set(
 echo = FALSE,
 fig.width = 7, 
 fig.height = 5,
 fig.asp = 0.6,
 out.width = "60%")
theme_set(theme_bw() + 
          theme(legend.position = "bottom",
                legend.title = element_blank(),
                plot.title = element_text(hjust = 0.5, size = 15),
                plot.subtitle = element_text(hjust = 0.5, size = 12)))
```

```{r, echo = FALSE, warning= FALSE, message=FALSE}
load_df =
GET("https://chronicdata.cdc.gov/resource/fqb7-mgjf.csv",
    query = list("$limit" = 50000)) %>% 
  content("parsed") %>% 
  janitor::clean_names()


ne_df = 
  load_df %>% 
  filter(is.na(data_value_footnote)) %>% 
  select(-class, -topic, -question, -data_value_unit, -data_value_type, -data_source, -data_value_footnote_symbol, -data_value_footnote, -question_id, -class_id, -topic_id, -states, -counties) %>% 
  filter(locationabbr == "NE")

```

```{r , echo = FALSE, warning= FALSE, message=FALSE}
time_df = 
  ne_df %>% 
  filter(response_id %in% c("RESP039", "RESP040") ) %>% 
  select(year, response_id, sample_size) %>% 
  group_by(year, response_id) %>% 
  mutate(count_obese = sum(sample_size)) %>% 
  select(-sample_size) %>% 
  distinct()

total_df = 
  ne_df %>%
  select(year, response_id, sample_size) %>% 
  group_by(year) %>% 
  mutate(count_total = sum(sample_size)) %>% 
  select(-sample_size) %>% 
  filter(response_id %in% c("RESP039", "RESP040") ) %>% 
  distinct() 

rate_time = 
  left_join(time_df, total_df,by = c("year" , "response_id")) %>% 
  mutate(
    prevalence = count_obese / count_total,
    rate_lable = case_when(
      response_id == "RESP039" ~ "Obese (BMI 30.0 - 99.8)",
      response_id == "RESP040" ~ "Overweight (BMI 25.0-29.9)"
    )
    ) %>%
  mutate(prevalence = round(prevalence, digits = 4)) 
```


## The prevalence of obese and overweight in Nebraska over time

```{r, echo = FALSE, warning= FALSE, message=FALSE, fig.align='center'}
rate_time %>%
  group_by(rate_lable) %>% 
  mutate(text_label = str_c("Year:", year, "\nPrevalence:", prevalence)) %>% 
  plot_ly( 
    x = ~year, 
    y = ~prevalence, 
    #type = 'scatter', 
    #mode = "lines+markers", 
    textposition = "bottom right",  
    color = ~rate_lable, 
    colors = c('#FFD700', '#FF8C00'),
    text = ~text_label, 
    line = list(simplyfy = F),
    alpha = 0.8
    ) %>% 
  layout(
    legend = list(x = 0.5, y = 0.15),
    title = "Obesity rate trend over time",
         xaxis = list(title = "Year",
                      zeroline = TRUE),
         yaxis = list(title = "Prevalence",
                      zeroline = TRUE),
     updatemenus = list(
      list(
        x = 0.7,
        y = 0.4,
        buttons = list(

          list(method = "restyle",
               args = list("type", "scatter"),
               label = "Line"),

          list(method = "restyle",
               args = list("type", "bar"),
               label = "bar")))
    )
    ) 


```

According to overview, Nebraska owns the highest rate of overweight and obese. From the line plot, we can see the prevalence of overweight decreased slightly while the prevalence of obese increased significantly over year. The total rate of people whose BMI is greater than 25 reaches almost 70% _(rate of obese is 36%)_ in 2019.

## Obesity by Race/Ethnicity

```{r, echo = FALSE, warning= FALSE, message=FALSE, fig.align='center'}
ne_race_df = 
  ne_df %>% 
  filter(break_out_category_id == "CAT4") %>% 
  select(year, locationabbr, locationdesc, response, break_out, sample_size, response_id) %>% 
  group_by(year, break_out) %>% 
  mutate(count_race = sum(sample_size)) %>% 
  filter(response_id %in% c("RESP039") ) %>% 
  mutate(race_pre = round(sample_size / count_race, digit = 4))

obesity_rate_by_race = 
  ne_race_df %>% 
  group_by(break_out) %>% 
  mutate(sum_race = sum(sample_size),
         sum_obese = sum(count_race),
         obesity_rate = round(sum_race/sum_obese, digits = 4)) %>% 
  select(break_out, obesity_rate) %>% 
  distinct() %>% 
  arrange(desc(obesity_rate))

fig_race_2 = 
  ne_race_df %>% 
  filter(break_out != "Other, non-Hispanic") %>% 
  mutate(text_label = str_c("Race/Ethnicity:", break_out, "\nPrevalence:", race_pre)) %>% 
  plot_ly( x = ~year , y = ~race_pre,  type = "bar", color = ~break_out, text = ~text_label, alpha = 0.7) %>% 
  layout(
         
         title = "Difference between race group over year",  
         legend = list(font = list(size = 8),  x = 0.5, y = 0.05),
         xaxis = list(title = "Year",
                      zeroline = TRUE,
                            rangeselector = list(
        buttons = list(
          list(
            count = 3,
            label = "3 year",
            step = "year",
            stepmode = "backward"),
          list(
            count = 6,
            label = "6 year",
            step = "year",
            stepmode = "backward"),
          list(
            count = 1,
            label = "YTD",
            step = "year",
            stepmode = "todate"),
          list(step = "all"))),

      rangeslider = list(type = "date")
                      ),
                      
         
         yaxis = list(title = "Prevalence",
                      zeroline = TRUE)) 


fig_race_2
```

`r obesity_rate_by_race %>% knitr::kable(col.names = c("Race/Ethnic", "Obesity rate"))`
Combining data from 2011 through 2019, _**non-Hispanic American Indian or Alaskan Native**_ had the highest prevalence of self-reported obesity (47.13%), followed by _**non-Hispanic Black adults**_ (41.81%) and 	
_**non-Hispanic Multiracial**_(39.74%). _**Non-Hispanic White adults**_ own the lowest obesity rate(31.95%).

## Women have lower obesity rate
```{r, echo = FALSE, warning= FALSE, message=FALSE, fig.align='center'}
ne_gender_df = 
  ne_df %>% 
  filter(break_out_category_id == "CAT2") %>% 
  select(year, locationabbr, locationdesc, response, break_out, sample_size, response_id) %>% 
  group_by(year, break_out) %>% 
  mutate(count_gender = sum(sample_size)) %>% 
  filter(response_id %in% c("RESP039") ) %>% 
  mutate(gender_pre = round(sample_size / count_gender, digit = 4))

ay = list(
  tickfont = list(color = "#767B84"),
  overlaying = "y",
  side = "right",
  title = "Prevalence",
  automargin = TRUE
)
ne_gender_df %>%
  group_by(break_out) %>% 
  mutate(
    text_label_1 = str_c("gender:", break_out, "\nCount:", sample_size),
    text_label_2 = str_c("gender:", break_out, "\nPrevalence:", gender_pre)     
    ) %>% 
  plot_ly( 
    x = ~year, 
    y = ~sample_size,  
    type = "bar", 
    color = ~break_out, 
    colors = c("#F1948A", "#85C1E9"),
    text = ~text_label_1, 
    alpha = 0.7) %>% 
    add_lines(x = ~year, y = ~gender_pre, color = ~break_out, text = ~text_label_2, yaxis = "y2") %>% 
  layout(
         title = "Obesity rate by gender",
         xaxis = list(title = "year",
                      zeroline = TRUE),
         yaxis = list(title = "Count",
                      zeroline = TRUE),
         yaxis2 = ay,
         legend = list(  
                     x = 0.7,
                     y = 0.1)
         )



```

According to the plot, the obesity rate of both women and men kept **increasing** over year and exceeded 0.35 since 2018. However, obesity rate of _**women**_ always significantly lower than that of _**man**_. 


## Obesity approximately decreased by level of education.

```{r, echo = FALSE, warning= FALSE, message=FALSE, fig.align='center'}
ne_edu_df = 
  ne_df %>% 
  filter(break_out_category_id == "CAT5") %>% 
  select(year, locationabbr, locationdesc, response, break_out, sample_size, response_id) %>% 
  group_by(year, break_out) %>% 
  mutate(count_edu = sum(sample_size)) %>% 
  filter(response_id %in% c("RESP039") ) %>% 
  mutate(edu_pre = round(sample_size / count_edu, digit = 4))

edu_rate = 
  ne_edu_df %>% 
  select(break_out, count_edu, year) %>% 
  distinct() %>% 
  group_by(break_out) %>% 
  mutate(sum_edu = round(sum(count_edu/9), digits = 0))

fig_edu_1 = 
  edu_rate %>% 
  plot_ly(labels = ~break_out, values = ~sum_edu, 
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#483C32'),
        hoverinfo = 'text',
        domain = list(row = 0, column = 1),
        text = ~paste('Education Attained:', break_out, ' \nNumber', count_edu),
        marker = list(colors = c('#6f4d96','#f3cec9','#cd7eaf','#e7a4b6'),
                      line = list(color = '#FFFFFF', width = 1)),
        showlegend = F
        ) %>% add_pie(hole = 0.5)

fig_edu_2 = 
  ne_edu_df %>% 
  group_by(break_out) %>% 
  mutate(text_label = str_c("Education Attained:", break_out, "\nPrevalence:", edu_pre)) %>% 
  plot_ly( 
    x = ~year , 
    y = ~edu_pre, 
    domain = list(row = 0, column = 0),
    type = 'scatter', 
    mode = "lines+markers", 
    #type = "bar", 
    color = ~break_out, 
    colors = c('#6f4d96', '#e7a4b6', '#f3cec9','#cd7eaf' ),
    text = ~text_label,
    alpha = 0.7) %>% 
  layout(
         title = "Obesity rate by Education",  
         legend = list( showlegend = T),
         xaxis = list(title = "Education Attained",
                      zeroline = TRUE),
         yaxis = list(title = "Prevalence",
                      zeroline = TRUE)) 

subplot(fig_edu_2 , fig_edu_1) %>%  layout(legend = list(x = 0.3, y = 0.04, font = list(size = 8), showlegend = F), grid = list(rows = 1, columns = 2))
```

-`College graduate`:College graduates

-`H.S. or G.E.D`: High school graduates or people pass General Educational Development test

-`Less than H.S`: People didn't attend high school

-`Some post-H.S`: Any school or specialized training which takes place after high school graduation

- Education level: `Less than H.S` < `H.S. or G.E.D`< `Some post-H.S` < `College graduate`.

The pie plot shows that the rate of **_College graduate_**, **_H.S or G.E.D_** and **_Some post-H.S_** are all around 30%, and the rate of **_Less than H.S_** is 6%. According to the line plot, the _obesity rate_ of all groups kept increasing over year. **_College graduate group_** always owns the _lowest_ obesity rate, which might benefit of their self-discipline. **_Less than H.S group_** owns the _highest_ obesity rate, however, the rate dropped dramatically between _2018_ and _2019_. The lines of **_Less than H.S group_**, **_Some post-H.S group_**, and **_H.S or G.E.D group_** intersects. From the trend, we assume that **Obesity approximately decreased by level of education**.
## Obesity rate trend of  income

```{r,fig.width = 15, fig.height = 11, out.width = "100%", echo = FALSE, warning= FALSE, message=FALSE, fig.align='center'}
ne_income_df = 
  ne_df %>% 
  filter(break_out_category_id == "CAT6") %>% 
  select(year, locationabbr, locationdesc, response, break_out, sample_size, response_id) %>% 
  group_by(year, break_out) %>% 
  mutate(count_income = sum(sample_size)) %>% 
  filter(response_id %in% c("RESP039", "RESP040")) %>% 
  mutate(
    income_pre = round(sample_size / count_income, digit = 4),
    rate_lable = case_when(
      response_id == "RESP039" ~ "Obese (BMI 30.0 - 99.8)",
      response_id == "RESP040" ~ "Overweight (BMI 25.0-29.9)")
    
    )
  

  ne_income_df %>% 
    mutate(break_out = factor(break_out, levels = c("Less than $15,000", "$15,000-$24,999", "$25,000-$34,999", "$35,000-$49,999", "$50,000+"))) %>% 
   ggplot(aes(x = year, y = income_pre, color = rate_lable)) +
   geom_line(size = 1) +
   scale_color_manual(values=c("#20B2AA", "#228B22")) +
   theme_bw() +
   theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45,  hjust = 1)) +
   facet_wrap(~break_out) +
   theme(legend.position = "bottom") +
    labs(title = "Obesity rate by household income",
       subtitle = "2011 - 2019.",
       x = "Year", 
       y = "Prevalence") +
  scale_x_continuous(breaks = seq(2011,2019,1)) 

```

From the plots, we can see people with lower household income tend to have higher obesity rate and lower overweight rate. The overweight rate decreased while obesity rate increased over time.